---
title: "Microrefugia-"
author: "Aji John"
date: "April 28, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE, message=F, warning=F}
#Packages
library("ggplot2")
load("/Users/ajijohn/git/climate/cr.RData")
```

## Site Description

#Shows the location of the Hobos
```{r,echo=FALSE, message=F, warning=F}
#load the raster, sp, and rgdal packages
library(raster)
library(sp)
library(rgdal)
library(leaflet)
library(jsonlite)
#lat,long
#38.579208	-122.700962
m <- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lng=-122.700962, lat=38.579208, popup="Pepperwood Center")
m
# adding custom tiles - Topo
m %>% addProviderTiles("Esri.WorldTopoMap")
# OpenTopoMap
m %>% addProviderTiles("OpenTopoMap")

# Esri.WorldImagery
#m %>% addProviderTiles("Esri.WorldImagery")

#Esri.NatGeoWorldMap
m %>% addProviderTiles("Esri.NatGeoWorldMap")

#Stamen.Terrain
m %>% addProviderTiles("Stamen.Terrain")

# Temp
#m %>% addProviderTiles("OpenWeatherMap.Temperature")

#NASAGIBS.ModisTerraChlorophyll
m %>% addProviderTiles("NASAGIBS.ModisTerraChlorophyll")

# add regions
geojson <- readLines("pw-regoins.json", warn = FALSE) %>%
  paste(collapse = "\n") %>%
  fromJSON(simplifyVector = FALSE)

# Default styles for all features
geojson$style = list(
  weight = 1,
  color = "#555555",
  opacity = 1,
  fillOpacity = 0.8
)
m %>% addGeoJSON(geojson)

outline <- ClimatePlotPts[chull(ClimatePlotPts$longitude, ClimatePlotPts$latitude),]

map <- leaflet(ClimatePlotPts) %>%
  # Base groups
  addTiles(group = "OSM (default)") %>%
  addProviderTiles("Esri.WorldTopoMap", group = "Esri Topo") %>%
  addProviderTiles("OpenTopoMap", group = "Open Topo") %>%
  addWMSTiles(
    "http://mesonet.agron.iastate.edu/cgi-bin/wms/nexrad/n0r.cgi",
    layers = "nexrad-n0r-900913",
    options = WMSTileOptions(format = "image/png", transparent = TRUE),
    attribution = "Weather data © 2012 IEM Nexrad"
    , group = "WMS") %>%
  # Overlay groups
  #addCircles(~Longitude, ~Latitude, ~10^1/5, stroke = F, group = "Quakes") %>%
  addPolygons(data = outline, lng = ~longitude, lat = ~latitude,
              fill = F, weight = 2, color = "#555555", group = "Outline") %>%
  
 # Layers control
  addLayersControl(
    baseGroups = c("OSM (default)", "Esri Topo", "Open Topo","WMS"),
    overlayGroups = c("Outline"),
    options = layersControlOptions(collapsed = FALSE)
  )
map
map%>% addCircleMarkers() %>%
addMarkers(~longitude, ~latitude, popup = ~as.character(Plot))
```

## Decoupling with respect to Santa Rosa(SR) - Further South West of Pepperwood

Santa Rosa is on a different elevation, much flatter terrain. To do a dry run, min/max of both the sites and the SR are compared with 'black' curve signifying the mean temperature of SR. 


#Trimmed the data to match the matrix i.e. matching days not found wrt to the site - Santa Rosa
```{r,echo=FALSE, message=F, warning=F}
library(weathermetrics)
library(R.oo)
plotList <- list()
plotList[]

for (i in 1:50){
 
  title = paste("Santa Rosa Matrix vs Hobo-" , hobo[i ,c("Plot")],sep="")
  # Make plot name
  pltName <- hobo[i ,c("Plot")]
  tmpHobomax = daily.max[[i]]
  tmpHobomin  = daily.min[[i]]
  tmpHobomax$dt = as.Date(paste(tmpHobomax$MONTH,'/',tmpHobomax$DAY,'/','14',sep =''), "%m/%d/%y")
  tmpHobomin$dt = as.Date(paste(tmpHobomin$MONTH,'/',tmpHobomin$DAY,'/','14',sep =''), "%m/%d/%y")
  #filter the Matrix to match 
  santarosa.2014.clean <- santarosa.qa.2014[match(tmpHobomax$dt,santarosa.qa.2014$date,),]
  santarosa.2014.clean$mean <- fahrenheit.to.celsius(santarosa.2014.clean$mean)
  santarosa.2014.clean$max <- fahrenheit.to.celsius(santarosa.2014.clean$max)
  santarosa.2014.clean$min <- fahrenheit.to.celsius(santarosa.2014.clean$min)

  tmpPlot = ggplot(data =santarosa.2014.clean, aes(x = date, y = mean)) +
  geom_line() +
  geom_line(mapping = aes(y = min),colour="green") +
  geom_line(mapping = aes(y = max),colour="red") + 
  geom_line(mapping = aes(y= tmpHobomin$x), colour="blue") + 
  geom_line(mapping = aes(y = tmpHobomax$x), colour="orange") + 
  ylab("Temp") + ggtitle(title)
  
 
  print(tmpPlot)     

  plotList[[pltName]] = tmpPlot

}

```

## Decoupling with respect to Windsor(WR)

Windsor again is on a different elevation, much flatter terrain. To do a dry run, min/max of both the sites and the WR are compared with 'black' curve signifying the mean temperature of SR. 

#Again, Trimmed the data to match the matrix i.e. matching days not found wrt to the site - Windsor.

```{r,echo=FALSE, message=F, warning=F}
library(weathermetrics)

plotList <- list()
plotList[]

for (i in 1:50){
 
  title = paste("Windsor Matrix vs Hobo-" , hobo[i ,c("Plot")],sep="")
  # Make plot name
  pltName <- hobo[i ,c("Plot")]
  tmpHobomax = daily.max[[i]]
  tmpHobomin  = daily.min[[i]]
  tmpHobomax$dt = as.Date(paste(tmpHobomax$MONTH,'/',tmpHobomax$DAY,'/','14',sep =''), "%m/%d/%y")
  tmpHobomin$dt = as.Date(paste(tmpHobomin$MONTH,'/',tmpHobomin$DAY,'/','14',sep =''), "%m/%d/%y")
  #filter the Matrix to match 
  windsor.2014.clean <- windsor.qa.2014[match(tmpHobomax$dt,windsor.qa.2014$date,),]
  
  tmpPlot = ggplot(data =windsor.2014.clean, aes(x = date, y = mean)) +
  geom_line() +
  geom_line(mapping = aes(y = min),colour="green") +
  geom_line(mapping = aes(y = max),colour="red") + 
  geom_line(mapping = aes(y= tmpHobomin$x), colour="blue") + 
  geom_line(mapping = aes(y = tmpHobomax$x), colour="orange") + 
  ylab("Temp") + ggtitle(title)
  
 
  print(tmpPlot)     

  plotList[[pltName]] = tmpPlot

}



```

## Minimum Temperature distribution across the sites.

#To use - Stability of min temperatures across the sites
#shows - very consistent minimum temperatures. Microrefugia sites tend to show stability, might be worthwhile to see it seasonally across years. We could see potentially increase in the minimum temperatures.

```{r,echo=FALSE, message=F, warning=F}
p1 <- ggplot(minTemps,aes(date,x,color=MONTH)) +
geom_point(data=minTemps[!(minTemps[,"site"]==""),],size=I(2),alpha=I(0.6)) 

pout <- p1 + facet_grid(. ~ site, scales = "free") +
xlab(paste("2014","")) + theme(axis.ticks = element_blank(),axis.text.x = element_blank()) 
#scale_color_manual(values=c("#666666", "#FF3300", "#0033FF" ))
pout
```

#Looking at the frequency of min temps across sites (bin size = 3 assuming we could see an increase of 3 deg)

```{r,echo=FALSE, message=F, warning=F}
ggplot(minTemps, aes(x, fill = site)) +
   geom_histogram(binwidth = 3)
```

## In Works

# detail daily trends
# elevation / slope / aspect study
```{r, echo=FALSE, message=F, warning=F}

```


